Sparse Matrix Vector Multiplication with Multiple vectors and Multiple Matrices on the MIC Architecture

### 2013-10-18 ieee_paper.tex
minor changes nad comments

We saw in the previous section that one can express the RBF problem as
a multiplication of four matrices by four vectors. We present here an
estimation of the variation on the flop intensity of the computation
and its impact on the expected performance of the
application. Relevant notations are given in Table~\ref{tab:not}.

\begin{figure}[bth]
  \begin{center}
    \scalebox{.8}{
      \begin{tabular}{|c|l|}
        %\hline
        %& & \\
        \hline
        $b_i$ & number of bytes per index \\
        $b_x$ & number of bytes per value \\
        $n_z$ & number of nonzeros per row of $A$ \\
        $n_r$ & number of column/rows of $A$ \\
        $n_c$ & total number of nonzeros\\
        $n_v$ & number of {\tt x} vectors \\
        $n_m$ & number of matrices \\
        $s_M$ & size of the $n_m$ matrices in bytes\\
        $s_x$ & size of the $n_v$ {\tt x} vectors in bytes\\
        $s_y$ & size of the $n_v n_m$ {\tt y} vectors in bytes\\
        \hline
        $cl$    & size of a cache line in bytes\\
        $b_{wT}$ & number of bytes written to memory  \\
        $b_{rT}$ & minimum number of bytes read from memory  \\
        $b_T$   & minimum number of bytes transferred \\
        $B_{rT}$ & maximum number of bytes read from memory \\
        $B_T$   & maximum number of bytes transferred \\
        $O$     & number of floating point operations \\
        $I_b$   & maximum computational intensity\\
        $I_w$   & minimum computational intensity\\        
        \hline
      \end{tabular}
    }
  \end{center}
  \caption{Notation relative to the matrices (upper section) and to the benchmarks (lower section.)}
  \label{tab:not}
\end{figure}

@@

%% In the special case $n_v=n_m=1$ in single precision
%% $$I_b = \frac{1}{4} \frac{ n_z }{ n_z + 1} \approx \frac{1}{4}$$
%% %$$I_w = \frac{n_z }{2  + 36  n_z} \approx \frac{1}{36}$$
%% $$I_w = \frac{n_z }{2  + (4+cl/2)  n_z} \approx \frac{1}{4+cl/2}$$

%% In the special case $n_v=n_m=4$ in single precision
%% $$I_b = \frac{32 n_z }{ n_z 20 + 80} \approx \frac{32}{20} = \frac{8}{5} $$
%% %$$I_w = \frac{32 n_z }{64 + n_z 84} \approx \frac{32}{84} = \frac{8}{21} $$
%% $$I_w = \frac{32 n_z }{64 + (20+cl) n_z} \approx \frac{32}{84} = \frac{8}{21} $$
%% The above formulas have been applied to MIC, which has a cache size $cl=64$ bytes.

%% \begin{table}
%%   \centering
%%   \begin{tabular}{|l|r|r|r|r|}
%%     \hline
%%                            & v1m1  & v1m4  & v4m1   & v4m4  \\
%%     \hline
%%     Best Single Precision  & 36.36 & 58.18 & 133.33 & 213.33\\
%%     Worst Single Precision &  4.15 & 14.20 &  16.55 &  55.81\\
%%     Best Double Precision  & 24.00 & 32.21 &  85.71 & 117.07\\
%%     Worst Double Precision &  3.93 & 11.88 &  15.58 &  46.15\\
%%     \hline
%%   \end{tabular}
%%   \caption{Estimation of performance in GFlop/s at 150GB/s for a
%%     matrix of $n_z = 32$}
%% \end{table}

@@

  \caption{Ratio of flops to bytes in single precision:  
    \ref{fig:ratio_best} best case; \ref{fig:ratio_worst} worst case when $cl=64$ 
    ; and \ref{fig:ratio_worst_nocache} worst case neglecting that the
    memory transfers are by batch (equivalent to $cl = 1$)}
  \label{fig:ratio_bytes_flops}
\end{figure*}

@@

\begin{figure}
  \centering 
  \includegraphics[width=.9\linewidth]{figures/gflops_peak.pdf}\label{fig:gflops-peak-perf}
  \caption{Estimation of the maximum achievable performance using a
    device with a memory to computational units of 150~GB/s when
    varying the number of vectors and matrices.}
  \label{fig:perf_predict}
\end{figure}

@@

  \caption{Bandwidth performance under idealized conditions.  Entries
    with "cpp" (triangles) denote cases where vectorization is
    compiler generated. \todo{legend} }
  \label{fig:bandwidth}
  \label{fig:band_rw_MIC}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\linewidth]{figures/bandwidth_gather_unpack.pdf}
  \caption{Performance of gather and unpack operations.}
  \label{fig:band_gather} 
\end{figure}
%

@@

  \caption{Performance of the manually vectorized code. The two
    horizontal lines depict the worst and the best predicted performances. }
  \label{fig:expe_types}
\end{figure*}

@@

  \caption{Performance of $y=Ax$ for a RBF derivative stencil of a
    $96^3$ 3D grid. Classical SpMV ($n_m=n_v=1$ given in blue) has low
    performance. The manually vectorized code (red) is 3.5 times
    faster than the compiler generated one (green). RCM ordering
    (squares) only improves the performance for our manually
    vectorized code.}
  \label{fig:comparison_spmv}
  \label{fig:perf_mic}
\end{figure}

>>>>

We saw in the previous section that one can express the RBF problem as
a multiplication of four matrices by four vectors. We present here an
estimation of the variation on the flop intensity of the computation
and its impact on the expected performance of the
application. Relevant notations are given in Figure~\ref{tab:not}.

\begin{figure}[bth]
  \begin{center}
    \scalebox{.8}{
      \begin{tabular}{|c|l|}
        %\hline
        %& & \\
        \hline
        $b_i$ & number of bytes per index \\
        $b_x$ & number of bytes per value \\
        $n_z$ & number of nonzeros per row of $A$ \\
        $n_r$ & number of column/rows of $A$ \\
        $n_c$ & total number of nonzeros\\
        $n_v$ & number of {\tt x} vectors \\
        $n_m$ & number of matrices \\
        $s_M$ & size of the $n_m$ matrices in bytes\\
        $s_x$ & size of the $n_v$ {\tt x} vectors in bytes\\
        $s_y$ & size of the $n_v n_m$ {\tt y} vectors in bytes\\
        \hline
        $cl$    & size of a cache line in bytes\\
        $b_{wT}$ & number of bytes written to memory  \\
        $b_{rT}$ & minimum number of bytes read from memory  \\
        $b_T$   & minimum number of bytes transferred \\
        $B_{rT}$ & maximum number of bytes read from memory \\
        $B_T$   & maximum number of bytes transferred \\
        $O$     & number of floating point operations \\
        $I_b$   & maximum computational intensity\\
        $I_w$   & minimum computational intensity\\        
        \hline
      \end{tabular}
    }
  \end{center}
  \caption{Notation relative to the application problem (upper section) and to the benchmarks (lower section.)}
  \label{tab:not}
\end{figure}

@@

@@

  \caption{Ratio of flops to bytes in single precision:  
    \ref{fig:ratio_best} best case; \ref{fig:ratio_worst} worst case when $cl=64$ 
    ; and \ref{fig:ratio_worst_nocache} worst case neglecting that the
    memory transfers are with a grenularity of one cache line (equivalent to $cl = 1$)}
  \label{fig:ratio_bytes_flops}
\end{figure*}

@@

\begin{figure}
  \centering 
  \includegraphics[width=.9\linewidth]{figures/gflops_peak.pdf}\label{fig:gflops-peak-perf}
  \caption{Estimation of the maximum achievable performance using a
    device with a memory to computational units of 150~GB/s when
    varying the number of vectors and matrices.\todo{yabel=Peak Performance (Gflop/s)}\todo{no xlabel}}
  \label{fig:perf_predict}
\end{figure}

@@

  \caption{Bandwidth performance under idealized conditions.  Entries
    with "cpp" (triangles) denote cases where vectorization is
    compiler generated. \todo{legend} \todo{ylabel=Bandwidth(GB/s)}}
  \label{fig:bandwidth}
  \label{fig:band_rw_MIC}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\linewidth]{figures/bandwidth_gather_unpack.pdf}
  \caption{Performance of gather and unpack operations.\todo{ylabel=Bandwidth(GB/s)}}
  \label{fig:band_gather} 
\end{figure}
%

@@

  \caption{Performance of the manually vectorized code. The two
    horizontal lines depict the worst and the best predicted performances. \todo{font size} \todo{ylabel=Performance(Gflop/s)}}
  \label{fig:expe_types}
\end{figure*}

@@

  \caption{Performance of $y=Ax$ for a RBF derivative stencil of a
    $96^3$ 3D grid. Classical SpMV ($n_m=n_v=1$ given in blue) has low
    performance. The manually vectorized code (red) is 3.5 times
    faster than the compiler generated one (green). RCM ordering
    (squares) only improves the performance for our manually
    vectorized code.\todo{ylabel=Performance(Gflop/s)}}
  \label{fig:comparison_spmv}
  \label{fig:perf_mic}
\end{figure}
